---
title: "DATA 621 – Business Analytics and Data Mining"
author: "Monu Chacko"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### In this homework assignment, you will explore, analyze and model a dataset containing approximately 8000 records representing  a  customer  at  an  auto  insurance  company. Each  record  has  two response variables.  The first responsevariable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero means that the person was not in a car crash.The second responsevariable is TARGET_AMT. This value is zero if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.


```{r message=FALSE}
library(e1071)
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(FactoMineR)
library(VIF)
library(knitr)
library(kableExtra)
library(Hmisc)
library(pROC)
library(binr)
library(MASS)
library(pscl)
library(AER)
```


```{r}
# read data
train = read.csv(file="data/wine-training-data.csv")
dim(train)
```