---
title: "Week 02 Assignment"
author: "Monu Chacko"
date: "February 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question

###Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.

###Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.

###This is by design a very open ended assignment.  A variety of reasonable approaches are acceptable.  You can (and should) blank out your SQL password if your solution requires it; otherwise, full credit requires that your code is “reproducible,” with the assumption that I have the same database server and R software.

###You may work in a small group on this assignment.   If you work in a group, each group member should indicate who they worked with, and all group members should individually submit their week 2 assignment.

###Please start early, and do work that you would want to include in a “presentations portfolio” that you might share in a job interview with a potential employer!  You are encouraged to share thoughts, ask, and answer clarifying questions in the “Week 2: R and SQL” forum.


###Answer:

####Create SQL table and store values. This data is stored in Azure SQL cloud DB.

```{r}
#IF OBJECT_ID('dbo.MovieReviews', 'U') IS NOT NULL  DROP TABLE dbo.MovieReviews; 

#CREATE TABLE [dbo].[MovieReviews]([ID] [int] IDENTITY(1,1) NOT NULL,	[Name] [varchar](150) NOT NULL,	[The_Cleanse] [int] NULL,	[Downsizing] [int] NULL,	[Lady_Bird] [int] NULL,	[Get_Out] [int] NULL,	[The_Shape_of_Water] [int] NULL,	[The_Post] [int] NULL, CONSTRAINT [PK_MovieReviews] PRIMARY KEY CLUSTERED (	[ID] ASC) WITH (STATISTICS_NORECOMPUTE = OFF, IGNORE_DUP_KEY = OFF) ON [PRIMARY]) ON [PRIMARY]

#INSERT INTO [dbo].[MovieReviews] ([Name] ,[The_Cleanse] ,[Downsizing] ,[Lady_Bird] ,[Get_Out] ,[The_Shape_of_Water] ,[The_Post]) VALUES  ('Carlos' ,1 ,4 ,5 ,3 ,3 ,2)

#INSERT INTO [dbo].[MovieReviews] ([Name] ,[The_Cleanse] ,[Downsizing] ,[Lady_Bird] ,[Get_Out] ,[The_Shape_of_Water] ,[The_Post]) VALUES  ('Victoria' ,3 ,4 ,2 ,1 ,2 ,3)

#INSERT INTO [dbo].[MovieReviews] ([Name] ,[The_Cleanse] ,[Downsizing] ,[Lady_Bird] ,[Get_Out] ,[The_Shape_of_Water] ,[The_Post]) VALUES  ('John' ,2 ,3 ,5 ,4 ,2 ,1)

#INSERT INTO [dbo].[MovieReviews] ([Name] ,[The_Cleanse] ,[Downsizing] ,[Lady_Bird] ,[Get_Out] ,[The_Shape_of_Water] ,[The_Post]) VALUES  ('Ahmed' ,5 ,2 ,5 ,3 ,1 ,3)

#INSERT INTO [dbo].[MovieReviews] ([Name] ,[The_Cleanse] ,[Downsizing] ,[Lady_Bird] ,[Get_Out] ,[The_Shape_of_Water] ,[The_Post]) VALUES  ('Paul' ,4 ,2 ,4 ,3 ,2 ,1)

#INSERT INTO [dbo].[MovieReviews] ([Name] ,[The_Cleanse] ,[Downsizing] ,[Lady_Bird] ,[Get_Out] ,[The_Shape_of_Water] ,[The_Post]) VALUES  ('Lou' ,5 ,4 ,4 ,3 ,2 ,4)

#SELECT ID, [Name] ,[The_Cleanse] ,[Downsizing] ,[Lady_Bird] ,[Get_Out] ,[The_Shape_of_Water] ,[The_Post] FROM [dbo].[MovieReviews]

```


####Read data from Cloud SQL DB 



```{r}
library(odbc)
library(getPass)
con <- dbConnect(odbc(),
                 Driver = "{ODBC Driver 13 for SQL Server}",
                 Server = getPass(msg="Server Name: "),
                 Database = getPass(msg="Database Name: "),
                 UID = getPass(msg="UID: "),
                 PWD = getPass(msg="Password: "),
                 Port = 1433)

#You can load using dns name
#con <- dbConnect(odbc::odbc(), "dnsname")

#Check if the table exist
dbListTables(con, table_name = "MovieR%")

#Read data
query<-paste0("SELECT ID, [Name] ,[The_Cleanse] ,[Downsizing] ,[Lady_Bird] ,[Get_Out] ,[The_Shape_of_Water] ,[The_Post] FROM [dbo].[MovieReviews]")
moviedata<-dbGetQuery(con,query)
moviedata

```


####Data cleanup

```{r}
#Drop ID
moviedata$ID <- NULL
moviedata
summary(moviedata)
```
