---
title: "Project 1"
author: "Monu Chacko"
date: "February 24, 2019"
output: html_document
---

###In this project, you’re given a text file with chess tournament results where the information has some structure.  Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

###Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

###For the first player, the information would be:

###Gary Hua, ON, 6.0, 1794, 1605

###1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played. 

##Answer:

###Lets load the content of the file and see the data

```{r message=FALSE, warning=FALSE}

library(stringr)

dschess <- readLines("./tournamentinfo.txt")
head(dschess)
tail(dschess)
```

###This data has to be cleaned up. We have to remove dashes. We can start by removing the header in the first 4 lines.  

```{r message=FALSE, warning=FALSE}
ds_cp_chess <- dschess[-c(0:4)]
head(ds_cp_chess, 20)
```

### Trim characters

```{r}
ds_cp_chess <- ds_cp_chess[sapply(ds_cp_chess, nchar) > 0]
```

###Extract line that contains rows with names into a variable. We can use seq() method to do this. This method returns row numbers from 1 to total length (192 rows) and skips by 3. Following are the rows that we will get.

```{r}
data_1 <- c(seq(1, length(ds_cp_chess), 3))
data_1
```

###Apply it to the dataset

```{r}
data_r1 <- ds_cp_chess[data_1]
head(data_r1)
```


###Extract name using regex

```{r}
name <- str_extract(data_r1, "[[:alpha:]]{2,}([[:blank:]][[:alpha:]]{1,}){1,}")
head(name)
```

###Extract the rows in the second row. Use the same technique as above.

```{r}
data_2 <- c(seq(2, length(ds_cp_chess), 3))
data_2
```

###Apply it to the dataset

```{r}
data_r2 <- ds_cp_chess[data_2]
head(data_r2)
```


###Extract state using regex

```{r}
state <- str_extract(data_r2, "[[:alpha:]]{2}")
state
```

###Extract points using regex

```{r}
pts <- str_extract(data_r1, "[[:digit:]]+\\.[[:digit:]]")
pts <- as.numeric(as.character(pts))
pts
```

###Extract pre rating using regex

```{r}
prertg <- str_extract(data_r2, ".\\: \\s?[[:digit:]]{3,4}")
prertg
```

###Extract digits using regex and convert it to numeric

```{r}
prertg <- as.numeric(str_extract(prertg, "\\(?[0-9,.]+\\)?"))
prertg

```

###Extract opponent number using regex. This data can be used to find opponents prerating average 

```{r message=FALSE, warning=FALSE}
oppnum <- str_extract_all(data_r1, "[[:digit:]]{1,2}\\|")
oppnum <- str_extract_all(oppnum, "[[:digit:]]{1,2}")
oppnum <- lapply(oppnum, as.numeric)
head(oppnum)
```

###Calculate prerating average for the opponent

```{r}
opppreratingavg <- list()

for (i in 1:length(oppnum)){
  opppreratingavg[i] <- round(mean(prertg[unlist(oppnum[i])]),2)
}
opppreratingavg <- lapply(opppreratingavg, as.numeric)
opppreratingavg <- data.frame(unlist(opppreratingavg))

df_final <- cbind.data.frame(name, state, pts, prertg, opppreratingavg)
colnames(df_final) <- c("Name", "State", "Points", "Pre_Rating", "Opp_Pre_Rating")
df_final
```

###Write the output to a file

```{r}
write.csv(df_final, "./ChessResults.csv")
```

###View stats

```{r message=FALSE, warning=FALSE}
library(ggplot2)

ggplot(df_final, aes(x=Pre_Rating)) + geom_histogram(binwidth = 50)

ggplot(df_final, aes(x=Opp_Pre_Rating)) + geom_histogram(binwidth = 50)

ggplot(data = df_final, aes(x = Pre_Rating, y = Opp_Pre_Rating)) + 
  geom_point(color='blue') +
  geom_smooth(method = "lm", se = FALSE)
```
