---
title: "Tidy Verse"
author: "Monu Chacko"
date: "5/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Data

NCAA Women's Basketball Tournament data contains information for every team that has participated in the NCAA Division I Women’s Basketball Tournament since it began in 1982. Every school is shown with its seed, conference record (when available), regular-season record, tournament record and full season record, including winning percentages. More information about select columns is in the table below.


```{r warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
library(tidyverse)
library(janitor)
library(kableExtra)
```

```{r}
dsWBB <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/ncaa-womens-basketball-tournament/ncaa-womens-basketball-tournament-history.csv", sep = ",", stringsAsFactors = F)

```


#### View raw data


```{r}
head(dsWBB) %>% kable() %>% kable_styling()

# Check columns
colnames(dsWBB)
```

#### Rename columns and view data

```{r}
# Rename columns
names(dsWBB) <- c("Year", "School", "Seed", "Conference", "ConfW", "ConfL","Conf", "ConfPlace", "RegW", "RegL", "Reg", "HowQual", "X1stGameAtHome", "TourneyW", "TourneyL", "TourneyFinish", "FullW", "FullL", "Full")

# Vew data
head(dsWBB) %>% kable() %>% kable_styling()

```


#### Drop columns that are not used and view data

```{r}
# Drop columns that are not necessary
dsWBB_new <- select(dsWBB, "Year", "Seed", "Conference", "HowQual", "X1stGameAtHome", "TourneyW", "TourneyL", "TourneyFinish", "FullW", "FullL", "Full")

# View new data
head(dsWBB_new) %>% kable() %>% kable_styling()

```


#### Extract and other data operations

```{r}
# Create new column from existing columns
dsWBB_new <- mutate(dsWBB_new, FullWToLRatio = round(dsWBB_new$FullW/dsWBB_new$FullL))

# Filter for year 1982
dsWBB_new_1982 <- filter(dsWBB_new, Year == "1982")

# View new data for 1982
head(dsWBB_new_1982) %>% kable() %>% kable_styling()

```


#### Sorting

```{r}

# Sort desc on FullWToLRatio
dsWBB_new_sorted <- arrange(dsWBB_new, desc(FullWToLRatio))

# Remove na
dsWBB_new_sorted$FullWToLRatio <- as.numeric(na.omit(dsWBB_new_sorted$FullWToLRatio))

# View sorted data
head(dsWBB_new_sorted) %>% kable() %>% kable_styling()
```


#### As we see in the above data, we have a divide by 0 error in column FullWToLRatio. We can replace all FullWToLRatio data with FullW where FullL is 0


```{r}
#Create new calculated colum
dsWBB_new_sorted = dsWBB_new_sorted %>% mutate(FullWToLRatio = factor(ifelse(FullL == 0, FullW, FullWToLRatio)))

# View sorted data
head(dsWBB_new_sorted) %>% kable() %>% kable_styling()

```


#### Splits and Levels

```{r}
# View Levels
FullWToLRatioLevels <- cut(as.numeric(dsWBB_new_sorted$FullWToLRatio), breaks = 10) %>% levels
FullWToLRatioLevels %>% kable() %>% kable_styling()
```


### Summarise and visualize data

```{r}

# Summarise
summarise(dsWBB_new_sorted, "MeanWLRatio" = mean(as.numeric(dsWBB_new_sorted$FullWToLRatio), na.rm = TRUE), "SDWLRatio" = sd(as.numeric(dsWBB_new_sorted$FullWToLRatio), na.rm = TRUE))

# View hist
hist(x=as.numeric(dsWBB_new_sorted$FullWToLRatio), main = "Win to Loss Ratio")
```

