---
title: "DATA 605 - Homework 10"
author: "Monu Chacko"
date: "4/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if 

### (a) he bets 1 dollar each time (timid strategy).

```{r}
q <- 0.6
p <- 0.4
s <- 1
M <- 8
qp <- q/p

qp

(qp^1-1)/(qp^8-1)
```

There is a 2% chance here.

### (b) he bets, each time, as much as possible but not more than necessary tobring his fortune up to 8 dollars (bold strategy).

```{r}
(0.4) ^ 3
```

OR

```{r}
dbinom(3,3,0.4)
```

There is a 6.4% chance with bold strategy.

### (c) Which strategy gives Smith the better chance of getting out of jail?


The bold strategy gives him better chance i.e 6.4% 

