---
title: "DATA 605 HW 12"
author: "Monu Chacko"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The attached who.csv dataset contains real-world data from 2008.  The variables included follow.

- Country:  name of the country
- LifeExp:  average life expectancy for the country in years
- InfantSurvival:  proportion of those surviving to one year or more
- Under5Survival:  proportion of those surviving to five years or more 
- TBFree:  proportion of the population without TB.
- PropMD:  proportion of the population who are MDs
- PropRN:  proportion of the population who are RNs
- PersExp:  mean personal expenditures on healthcare in US dollars at average exchange rate 
- GovtExp:  mean government expenditures per capita on healthcare, US dollars at average exchange rate
- TotExp:  sum of personal and government expenditures.


```{r}
library(knitr)
whodf <- read.csv(file="https://raw.githubusercontent.com/monuchacko/cuny_msds/master/data_605/who.csv", header=TRUE, sep=",")

kable(head(whodf), digits = 2, align = c(rep("l", 4), rep("c", 4), rep("r", 4)))
```


1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression.  Do not transform the variables.  Provide and interpret the F statistics, R^2, standard error,and p-values only.  Discuss whether the assumptions of simple linear regression met.2.Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6).  

```{r}
plot(whodf$LifeExp ~ whodf$TotExp, main = "LifeExp vs TotExp", xlab = "Pers and gov expenditures", ylab = "Average life expectancy")
abline(lm(whodf$LifeExp ~ whodf$TotExp), col="red") # regression line (y~x) 
```

```{r}
# linear regression
m1 <- lm(LifeExp ~ TotExp, data = whodf)
summary(m1)
```


From the data above we can say that there is high likelihood that the model is explaining the data failrly well. Here F-Stat is 65.26 and p-value is close to 0.

```{r}
qqnorm(m1$residuals)
qqline(m1$residuals)
```


2. Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6  as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables.  Provide and interpret the F statistics, R^2, standard error, and p-values.   Which model is "better?"

```{r}
whodf2<-whodf
whodf2$LifeExp<-whodf2$LifeExp^4.6
whodf2$TotExp<-whodf2$TotExp^0.6

kable(head(whodf2), digits = 2, align = c(rep("l", 4), rep("c", 4), rep("r", 4)))
```

```{r}
plot(whodf2$LifeExp ~ whodf2$TotExp, main = "LifeExpTransformed vs TotExpTransformed", xlab = "Pers and gov expenditures", ylab = "Average life expectancy")
abline(lm(whodf2$LifeExp ~ whodf2$TotExp), col="red") # regression line (y~x) 
```


```{r}
m2 <- lm(LifeExp ~ TotExp, data = whodf2)
summary(m2)
```

3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5.  Then forecast life expectancy when TotExp^.06=2.5.

```{r}
TExp <- 1.5
LExp <- 238461*TExp + 211907647
round(LExp ^ (1/4.6),1)

TExp <- 2.5
LExp <- 238461*TExp + 211907647
round(LExp ^ (1/4.6),1)
```


4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?

LifeExp = b0+b1 x PropMd + b2 x TotExp  +b3 x PropMD x TotExp

```{r}
m3 <- lm(LifeExp ~ PropMD + TotExp + PropMD*TotExp, data = whodf)
summary(m3)

qqnorm(m3$residuals)
qqline(m3$residuals)
```

5. Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?

```{r}
TExp <- 14
PrMD <- 0.03
LExp <- 6.277e+01 + 1.497e+03*PrMD + 7.233e-05*TExp -6.026e-03*PrMD*TExp
round(LExp,1)
```
