---
title: "DATA 605 Home Work 13"
author: "Monu Chacko"
date: "5/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1

Use integration by substitution to solve the integral below.

$$ \int{4e^{-7x}dx}$$

$$
u=-7x,\text{    }u'=-7x\ dx = -\frac{1}{7}\\
\int{4e^{-7x}}dx = 4 \int{e^u} du * u'\\
-\frac{4}{7}\int{e^u du}\\
-\frac{4}{7} e^u\\
-\frac{4}{7} e^{-7X}
$$


### Question 2

Biologists are treating a pond contaminated with bacteria.  The level of contamination is changing at a rate of $\frac{dN}{dt} = -\frac{3150}{t^4} - 220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began.  Find a function N(t) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.


Constant of integration, N0 if N(1) = 6530


$$\frac{dN}{dt}=-\frac{3150}{t^4} - 220$$

$$dN =(\frac{3150}{t^4}−220)dt$$

$$N=\int{\frac{3150}{t^4}dt} −\int{220dt}$$

$$N = N_0 − \frac{3150}{3t^3}− 220t$$

$$N(_1) = N_0 −\frac{1050}{1^3}−220(1)$$

$$N_0 = 6530 + 1050 + 220$$

$$N_0 = 7800$$

$$N = 7800−\frac{1050}{t^3}−220t$$

### Question 3

Find the total area of the red rectangles in the figure below, where the equation of the line is $f(x) = 2x-9$.


Area is a triangle from 4.5 to 8.5.

```{r }
funcArea <- function(x){
  2*x - 9
}
integrate(funcArea, 4.5, 8.5)
```

### Question 4

Find the area of the region bounded by the graphs of the given equations.

$$
\begin{split}
y &= x^2-2x-2 \\ 
 y &= x+2
\end{split}
$$


```{r}
X = function(x) {x + 2}
Y = function(x) {x^2 -2*x -2}
```

```{r}
ax <- integrate(X, -1, 4)
yx <- integrate(Y, -1, 4)
(Area <- round((ax$value - yx$value),4))
```

### Question 5

A beauty supply store expects to sell 110 flat irons during the next year.  It costs $3.75 to store one flat iron for one year.  There is a fixed cost of $8.25 for each order.  Find the lot size and the number of orders per year that will minimize inventory costs.


$$Storage\ cost/ Yr = 3.75 \times x/2 = 1.875x$$

$$Ordering\ cost/Yr = 8.25 \times 110/x = 907.5/x$$

$$Inventory\ cost = 1.875x+907.5/x = f(x)$$



$$
\begin{split}
f'(x) &= 1.875-\frac{907.5}{x^2} \\
f'(x) &= 0 
\end{split}
$$


$$
\begin{split}
1.875-\frac{907.5}{x^2} &= 0 \\
1.875&= \frac{907.5}{x^2} \\
1.875x^2&= 907.5 \\
x^2&= \frac{907.5}{1.875} \\
x&= \sqrt{\frac{907.5}{1.875}} \\
x&=\sqrt{484} \\
x&=22
\end{split}
$$
$$ orders = 110/22=5$$

### Question 6

Use integration by parts to solve the integral below.

$$ \int { ln(9x).{ x }^{ 6 }dx }  $$


$$=\frac{1}{7}x^7 * ln(9x) − \int{\frac{1}{7}x^7 * \frac{1}{x}dx}$$

$$=\frac{1}{7}x^7 * ln(9x) - \int{\frac{1}{7}x^6dx}$$

$$= \frac{7}{49}x^7 * in(9x) - \frac{1}{49} x^7dx + C$$

$$= \frac{1}{49}x^7*(7ln(9x) - 1) + C$$



### Question 7

Determine whether f (x) is a probability density function on the interval 1, e6.  If not, determine the value of the definite integral.

$$f(x) = \frac{1}{6x}$$


Is the area under the curve within the interval equal to 1 

$$F(x) = \int_1^{e^6}{\frac{1}{6x}dx}=1$$

$$
\begin{split}
\int_1^{e^6}{\frac{1}{6x}dx} &= \frac{1}{6}ln(x) |_1^{e^6} \\
&= \frac{1}{6}ln(e^6) - \frac{1}{6}ln(1) \\
&=\frac{1}{6}\times6 - \frac{1}{6}\times 0 \\
&= 1 - 0 \\
&= 1
\end{split}
$$

f(x) is a probability density function
