---
title: "Data 605 - Assignment 15"
author: "Monu Chacko"
date: "5/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Find the equation of the regression line for the given points.  Round any final values to the nearest hundredth, if necessary.

#### ( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

```{r}
df <- data.frame(X = c(5.6, 6.3, 7, 7.7, 8.4), Y = c(8.8, 12.4, 14.8, 18.2, 20.8))
md1 <- lm(Y ~ X, data = df)
s1 <- summary(md1)
s1
```

y = 4.26x -14.80


### 2. Find all local maxima, local minima, and saddle points for the function given below.  Write your answer(s) in the form (x, y, z).  Separate multiple points with a comma.

$$f\left( x,y \right) =24x-6x{ y }^{ 2 }-8{ y }^{ 3 }$$

$$\frac { df }{ dx } =24-6{ y }^{ 2 }$$

$$\frac { df }{ dy } =-12xy-24{ y }^{ 2 }$$

$$if\_ y\Rightarrow 0=24-6{ y }^{ 2 }\Rightarrow { y }^{ 2 }=4\Rightarrow y\pm 2$$

$$if\_ y\Rightarrow \pm 2...\& ...-12xy-24{ y }^{ 2 }=0\Rightarrow \pm { 24 }x=24x4\Rightarrow x=\pm 4$$




### 3. A grocery store sells two brands of a product, the "house" brand and a "name" brand.  The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the "house" brand and 40 + 11x - 23y units of the "name" brand.

#### Step 1. Find the revenue function R (x, y).

$$R(x,y) = (81 - 21x + 17y)x + (40 + 11x - 23y)y$$

$$R(x,y) = 81x - 21x^2 + 17yx + 40y + 11xy - 23y^2$$
$$R(x,y) = 81x - 21x^2 + 28yx + 40y - 23y^2$$


#### Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

```{r}
x=2.30
y=4.10

s2 <- -21*x^2 +81*x +28*x*y +40*y -23*y^2
s2

```

### 4. A company has a plant in Los Angeles and a plant in Denver.  The firm is committed to produce a total of 96 units of a product each week.  The total weekly cost is given by $C\left( x,y \right) =\frac { 1 }{ 6 } { x }^{ 2 }+\frac { 1 }{ 6 } { y }^{ 2 }+7x+25y+700$ where x is the number of units produced in Los Angeles and y is the number of units produced in Denver.  How many units should be produced in each plant to minimize the total weekly cost?

$$
\begin{align} 
C(x,y) & = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700 & \text{ } \\
  f(x) & = \frac{1}{6}x^2 + \frac{1}{6}(96-x)^2 + 7x + 25(96-x) + 700 &  \text{substitute y=96-x} \\
  f'(x)& = \frac{1}{6}2x  + \frac{1}{6}2(96-x)(-1) + 7 + 25(-1) & \\
  f'(x)& = \frac{1}{3}x   - \frac{1}{3}(96-x) - 18 & \\
  f'(x)& = \frac{2}{3}x - 32 - 18 & \\
  f'(x)& = \frac{2}{3}x - 50 & \\
  f''(x) &= \frac{2}{3} > 0
\end{align}
$$

$f'(x) = 0$ -> $x = 75$  


Values of x:

$$0 \leq x \leq 96$$

$$y= 96-x = 21$$

$$y = 75; y = 21$$


### 5. Evaluate the double integral on the given region.

$$\int \int_{R} e^{8x + 3y} dA  \text{ where }  R:=\bigg \lbrace  2 \leq x \leq 4;  2 \leq y \leq 4\bigg \rbrace$$

### Write your answer in exact form without decimals.


$$
\begin{align}
\int \int_{R} e^{8x + 3y} dA & = & \bigg(\int_{2}^{4} e^{8x}dx \bigg)\bigg( \int_{2}^{4}e^{3y}dy \bigg) \\
    & = & \bigg(  \frac{1}{8}e^{8x} \bigg \lvert_{x=2}^{4}\bigg) \bigg(  \frac{1}{3} e^{3y} \bigg \lvert_{y=2}^{4} \bigg) \\
    & = & \bigg( \frac{ e^{32} - e^{16}}{8}  \bigg) \bigg( \frac{ e^{12} - e^{6}  }{ 3}   \bigg) \\
    & = & \frac{1}{24}(e^{32}-e^{16})(e^{12}- e^{6})
\end{align}
$$