---
title: "DATA 605 HW 2"
author: "Monu Chacko"
date: "2/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS

> Problem Set 1

(1) Show that A<sup>T</sup> A != AA<sup>T</sup> in general. (Proof and demonstration.)

#### Answer 1  

**Proof with out calculation**  

The Commutative property of multiplication is not always true in a matrix multiplication. Order of multiplication matters. Lets say A<sup>T</sup> is a 5x2 matrix and A is a 2x3 matrix then its can be multiplied because the A<sup>T</sup> has the same number of columns as the number of rows for A i.e 2. Its product will yeild a 5x3 matrix.

So let say we change the order and multiply i.e AA<sup>T</sup> then it becomes *undefined because the number of columns for A will not be same as the number of rows for A<sup>T</sup>*.  

**Proof with calculation** 

Lets take another example, this time where both are not undefined. 

$$
A^T =
\left[
  \begin{array}{cccc}
  3 & 4 \\ 2 & 1
  \end{array}
\right]
 A =
\left[
  \begin{array}{cccc}
  3 & -4 \\ 1 & 2
  \end{array}
\right]
$$

Find A<sup>T</sup> A  

$$
A^T A =
\left[
  \begin{array}{cccc}
  3 & 4 \\ 2 & 1
  \end{array}
\right]
\left[
  \begin{array}{cccc}
  3 & -4 \\ 1 & 2
  \end{array}
\right]
$$


$$
A^T A =
\left[
  \begin{array}{cccc}
  13 & -4 \\ 7 & -6
  \end{array}
\right]
$$


Let us now find AA<sup>T</sup>  


$$
A A^T =
\left[
  \begin{array}{cccc}
  3 & -4 \\ 1 & 2
  \end{array}
\right]
\left[
  \begin{array}{cccc}
  3 & 4 \\ 2 & 1
  \end{array}
\right]
$$

$$
A A^T =
\left[
  \begin{array}{cccc}
  1 & 7 \\ 8 & 6
  \end{array}
\right]
$$

--------

(2) For a special type of square matrix A, we get A<sup>T</sup> A = AA<sup>T</sup>. Under what conditions could this be true?  (Hint:  The Identity matrix I is an example of such a matrix). Please type set your response using LaTeX mode in RStudio. 

#### Answer 2  

The Commutative property of multiplication is true for an identity matrix. For example:

$$
IA =
\left[
  \begin{array}{cccc}
  1 & 0 \\ 0 & 1
  \end{array}
\right]
\left[
  \begin{array}{cccc}
  5 & 3 \\ 3 & 5
  \end{array}
\right]
=
\left[
  \begin{array}{cccc}
  5 & 3 \\ 3 & 5
  \end{array}
\right]
$$

$$
AI =
\left[
  \begin{array}{cccc}
  5 & 3 \\ 3 & 5
  \end{array}
\right]
\left[
  \begin{array}{cccc}
  1 & 0 \\ 0 & 1
  \end{array}
\right]
=
\left[
  \begin{array}{cccc}
  5 & 3 \\ 3 & 5
  \end{array}
\right]
$$

Another example:

$$
IA =
\left[
  \begin{array}{cccc}
  1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1
  \end{array}
\right]
\left[
  \begin{array}{cccc}
  5 & 3 & 2 \\ 3 & 5 & 9 \\ 2 & 7 & 7
  \end{array}
\right]
=
\left[
  \begin{array}{cccc}
  5 & 3 & 2 \\ 3 & 5 & 9 \\ 2 & 7 & 7
  \end{array}
\right]
$$

$$
AI =
\left[
  \begin{array}{cccc}
  5 & 3 & 2 \\ 3 & 5 & 9 \\ 2 & 7 & 7
  \end{array}
\right]
\left[
  \begin{array}{cccc}
  1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1
  \end{array}
\right]
=
\left[
  \begin{array}{cccc}
  5 & 3 & 2 \\ 3 & 5 & 9 \\ 2 & 7 & 7
  \end{array}
\right]
$$

--------

> Problem Set 2

Matrix factorization is a very important problem.  There are super computers built just to  do  matrix  factorizations.   Every  second  you  are  on  an  airplane,  matrices  are  being factorized.  Radars that track flights use a technique called Kalman filtering.  At the heart of Kalman Filtering is a Matrix Factorization operation.  Kalman Filters are solving linear systems of equations when they track your flight using radars. Write an R function to factorize a square matrix A into LU or LDU, which ever you prefer. You don’t have to worry about permuting rows of A and you can assume that A is lessthan  5x5,  if  you  need  to  hard-code  any  variables  in  your  code.   If  you  doing  the  entire assignment in R, then please submit only one markdown document for both the problems.

> LU Decomposition - Shortcut Method  

```{r warning=FALSE}
luDecompose <- function(A) {
  # Make sure A is a square matrix
  if (dim(A)[1]!=dim(A)[2]) {
    print("Err: A is not a square matrix")
    return(NA)
  }
  
  U <- A
  n <- dim(A)[1]
  L <- diag(n)
  
  # If dimension is 1, then return matrix
  if (n==1) {
    return(list(L,U))
  }
  
  # Loop through the lower triangle and determine multiplier for each position and add it to L
  for(i in 2:n) {
    for(j in 1:(i-1)) {
      multiplier <- -U[i,j] / U[j,j]
      U[i, ] <- multiplier * U[j, ] + U[i, ]
      L[i,j] <- -multiplier

      print("=========U=========")
      print(U)
      print("=========L=========")
      print(L)
    }
  }
  return(list(L,U))
}
```


#### Test single 1x1 matrix  

```{r warning=FALSE}
A <- matrix(c(1), nrow=1, byrow=TRUE)
LU <- luDecompose(A)
if (!is.na(LU)) {
  L<-LU[[1]]  
  U<-LU[[2]]
  A
}
```

#### Test rectangle matrix
This should return error  

```{r warning=FALSE}
A = matrix(c(2, 4, 3, 1, 5, 7), nrow=2, ncol=3, byrow = TRUE)
LU <- luDecompose(A)
if (!is.na(LU)) {
  L<-LU[[1]]  
  U<-LU[[2]]
  A
}
```

#### Test 4x4 square matrix  

```{r warning=FALSE}
A <- matrix(c(11,12,-4,5,6,9,7,6,4,3,2,5,2,1,4,2), nrow=4, byrow=TRUE)
LU <- luDecompose(A)
if (!is.na(LU)) {
  L<-LU[[1]]  
  U<-LU[[2]]
  A
}
```

#### Verify the test for square matrix  

```{r}
A == L %*% U
```
