---
title: "FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS"
author: "Monu Chacko"
date: "02/23/2020"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ASSIGNMENT 5


Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions.Note that the point (B,C) is then chosen at random in the unit square.Find the probability that  

Uniform Probability Distribution

### Prepare data  

```{r}
n <- 1000

# Default min - 0 and max - 1
B = runif(n, min = 0, max = 1)
C = runif(n, min = 0, max = 1)
```


### Check data


```{r}
plot(1:n, B)
plot(1:n, C)

hist(B)
hist(C)
```


### (a) B + C < 1/2.

Solution using custom function

```{r}
s <- function(B, C) {
  val <- 0
  cn <- 0
  total_len <- length(B)
  for(i in 1:length(B)) {
    val <- B[i] + C[i]
    if (val < .5) {
      cn = cn + 1    
    }
  }
  
  p <- cn / total_len
  return(p)
}


s(B, C)
```


Solution using built-in function

```{r}
D <- B+C
sum(punif(D<1/2, min=0, max = 1))/n
```


### (b) BC < 1/2.

Solution using custom function

```{r}
s <- function(B, C) {
  val <- 0
  cn <- 0
  total_len <- length(B)
  for(i in 1:length(B)) {
    val <- B[i] * C[i]
    if (val < .5) {
      cn = cn + 1    
    }
  }
  
  p <- cn / total_len
  return(p)
}

s(B, C)
```


Solution using built-in function

```{r}
E <- B*C
sum(punif(E<1/2, min=0, max = 1))/n
```


### (c) |B − C| < 1/2.

Solution using custom function

```{r}
s <- function(B, C) {
  val <- 0
  cn <- 0
  total_len <- length(B)
  for(i in 1:length(B)) {
    val <- B[i] - C[i]
    if (abs(val) < .5) {
      cn = cn + 1    
    }
  }
  
  p <- cn / total_len
  return(p)
}

s(B, C)

```

Solution using built-in function

```{r}
G <- abs(B-C)
sum(punif(G<1/2, min=0, max = 1))/n
```


### (d) max{B,C} < 1/2.

Solution using custom function

```{r}
s <- function(B, C) {
  val <- 0
  cn <- 0
  total_len <- length(B)
  for(i in 1:length(B)) {
    val <- pmax(B[i],C[i])
    if (val < .5) {
      cn = cn + 1    
    }
  }
  
  p <- cn / total_len
  return(p)
}

s(B, C)
```


Solution using built-in function

```{r}
H <- pmax(B,C)
sum(punif(H<1/2, min=0, max = 1))/n
```


### (e) min{B,C} < 1/2.

Solution using custom function

```{r}
s <- function(B, C) {
  val <- 0
  cn <- 0
  total_len <- length(B)
  for(i in 1:length(B)) {
    val <- pmin(B[i],C[i])
    if (val < .5) {
      cn = cn + 1    
    }
  }
  
  p <- cn / total_len
  return(p)
}

s(B, C)
```


Solution using built-in function

```{r}
H <- pmin(B,C)
sum(punif(H<1/2, min=0, max = 1))/n
```

