---
title: "DATA 605 HW 1"
author: "Monu Chacko"
date: "02/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### PROBLEM SET 1

1. Calculate the dot product u.v where u = [0.5; 0.5] and v = [3;−4]

```{r}
library(geometry)
u <- c(0.5, 0.5)
v <- c(3, -4)
u.v <- dot(u,v,d = NULL)
u.v
```

$$
u =
\left[
  \begin{array}{cccc}
  0.5 & 0.5
  \end{array}
\right]
 v =
\left[
  \begin{array}{cccc}
  3 & -4
  \end{array}
\right]
$$

$$
u.v = (0.5 \times 3) + (0.5 \times -4) = -0.5
$$

2. What are the lengths of u and v?  Please note that the mathematical notion of the length of a vector is not the same as a computer science definition.

```{r}
len_u <- sqrt(sum(u^2))
len_u

len_v <- sqrt(sum(v^2))
len_v

```

Solving using pythagorean theorem:

$$
||\overrightarrow{u}|| = \sqrt{0.5^2 + 0.5^2} = 0.7071 
$$
$$
||\overrightarrow{v}|| = \sqrt{3^2 + -4^2} = 5 
$$
3. What is the linear combination:  3u−2v?


```{r}
lc <- 3*u - 2*v
lc
```


$$
\overrightarrow{u} = \left[
  \begin{array}{cccc}
  0.5 & 0.5
  \end{array}
\right]
$$

$$
3\overrightarrow{u} = \left[
  \begin{array}{cccc}
  1.5 & 1.5
  \end{array}
\right]
$$
$$
\overrightarrow{v} = \left[
  \begin{array}{cccc}
  3 & -4
  \end{array}
\right]
$$


$$
2\overrightarrow{v} = \left[
  \begin{array}{cccc}
  6 & -8
  \end{array}
\right]
$$

$$
3\overrightarrow{u} - 2\overrightarrow{v} =
\left[
  \begin{array}{cccc}
  1.5 & 1.5
  \end{array}
\right] -
\left[
  \begin{array}{cccc}
  6 & -8
  \end{array}
\right]
$$

$$
(1.5 - 6) = -4.5
$$
$$
(1.5 + 8) = 9.5
$$

$$
3\overrightarrow{u} - 2\overrightarrow{v} = 
\left[
  \begin{array}{cccc}
  -4.5 & 9.5
  \end{array}
\right]
$$

4. What is the angle between u and v

```{r}
angle <- acos( dot(u, v) / ( sqrt(dot(u, u)) * sqrt(dot(v, v)) ) )
angle_degree <- angle * 180 / pi
angle_degree
```


### PROBLEM SET 2

Set up a system of equations with 3 variables and 3 constraints and solve for x.  Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination.  Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system.  You don’t have to worry about degenerate cases and can safely assume that the function will only be tested with a system of equations that has a solution. Please note that you do have to worry about zero pivots, though. Please note that you should not use the built-in function solve to solve this system or use matrix inverses.  The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution.  Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure. 

Please test it with the system below and it should produce a solution x = [−1.55,−0.32,0.95]

$$
3\overrightarrow{u} - 2\overrightarrow{v} = 
\left[
  \begin{array}{cccc}
  1 & 1 & 3 \\
  2 & -1 & 5 \\
  -1 & -2 & 4 \\
  \end{array}
\right]
\left[
  \begin{array}{cccc}
  x1 \\
  x2 \\
  x3 \\
  \end{array}
\right]
=
\left[
  \begin{array}{cccc}
  1 \\
  2 \\
  6 \\
  \end{array}
\right]
$$


```{r}
# Function to check for upper triangular matrix
is_upper_triangular <- function(A) {
  is_ut_matrix = TRUE
  
  for(row in 1:nrow(A)) {
    for(col in 1:ncol(A)) {
      # If row# is greater than col# then they should be 0 (below the main diagonal) 
      if (row > col) {
        if (A[row, col] != 0) {
          is_ut_matrix = FALSE
        }
      }
    }
  }
  return(is_ut_matrix)
}
```


```{r}
func_sol_set2 <- function(A, b)
{
  A <- cbind(A, b)
  m <- A[2,1]/A[1,1]*(A[1,])
  A[2,] = A[2,]-m

  print("Showing steps of operation")
  print("Step 1")  
  print(A)
  print(" ")
  
  m = (A[3,1]/A[1,1]*(A[1,]))
  A[3,] = A[3,]-m
  
  print("Step 2")  
  print(A)
  print(" ")
  
  m = (A[3,2]/A[2,2]*(A[2,]))
  A[3,] = A[3,]-m

  print("Step 3")  
  print(A)
  print(" ")
  print("Upper Triangular Check")
  print(is_upper_triangular(A))
  
  x3 <- (A[3,4]/A[3,3])
  x2 = (A[2,4]-A[2,3]*x3)/A[2,2]
  x1 = (A[1,4]-A[1,3]*x3 - A[1,2]*x2)/A[1,1]
  
  return(c(x1, x2, x3))
}

A <-  matrix(
  c(1, 2, -1, 1, -1, -2, 3, 5, 4), 
  nrow=3, 
  ncol=3)

b <- matrix(
  c(1, 2, 6),
  nrow = 3,
  ncol = 1)

```


```{r}
# Tests
print(is_upper_triangular(A))
r <- func_sol_set2(A, b)

# Show results
print(r)
```



