---
title: "Data 605 HW7"
author: "Monu Chacko"
date: "3/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### 1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi's. Find the distribution of Y.

If $K^n$ is the sum of variables then $(k-1)^n$ would represent varaibles where $X_i$ does not contain 1

$$P(X=1)=\frac { { k }^{ n }-{ (k-1) }^{ n } }{ { k }^{ n } }$$

$$P(X=2)=\frac { { (k-2+1) }^{ n }-{ (k-2) }^{ n } }{ { k }^{ n } }$$

$$P(X=y)=\frac { { (k-y+1) }^{ n }-{ (k-y) }^{ n } }{ { k }^{ n } }$$


### 2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).  This machine has a manufacturer's expected lifetime of 10 years.  This means that we expect one failure every ten years.  (Include the probability statements and R Code for each part.). 

### a.  What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a geometric.  (Hint: the probability is equivalent to not failing during the first 8 years..) 

```{r}
prob_fail <- 1/10
prob_fail
prob <- ((1 - prob_fail)^ (8 - 1) * prob_fail)
prob

exp_value <- 1/prob_fail
exp_value

sd <- sqrt((1-prob_fail)/(prob_fail^2))
sd
```

### b.  What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as an exponential.

```{r}
l <- (1 / 10)
eprob <- exp(1) ^ (- l * 8)
eprob

evalue <- 1 / l
evalue

sd <- 1 / (l ^ 2)
sd
```

### c.  What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a binomial.  (Hint:  0 success in 8 years)

```{r}
p <- 1 / 10
yrs <- 8
suc <- 0

# Probability of failure
prob_b <- choose(yrs, suc)*((p)^(suc))*(1-p)^yrs
prob_b

# Probability of non failure
prob_not_b <- 1 - prob_b
prob_not_b

# Expected Value
exp_v <- yrs *p
exp_v

# Standard Deviation
sd <- sqrt(yrs*p*(1 - p))
sd
```

### d.  What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a Poisson.

$$\frac { { e }^{ -\lambda }{ \lambda }^{ n } }{ n! }$$

```{r}
yrs <- 8

lp = yrs * (1 / 10)
e <- exp(1)
prob_poi <-(e ^ (- lp) * lp ^ 0) / factorial(0)
prob_poi

pv_exp <- lp
pv_exp

sd <- sqrt(lp)
sd

paste("Failure probabiity after,", yrs, " years is", round(prob_poi, 2), ". The expected value is", pv_exp, ". The standard deviation is", round(sd, 2), sep = " ")
```
