---
title: "Assignment 9"
author: "Monu Chacko"
date: "3/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Q 11 (Page 363)

The price of one share of stock in the Pilsdorff Beer Company is given by $Y_n$ on the nth day of the year. Finn observes that the differences $X_n = Y_{n+1} - Y_n$ appear to be independent random variables with a common distribution having mean $\mu = 0$ and variance $\sigma^2 = 1/4$. If $Y_1 = 100$, estimate the probability that $Y_{365}$ is

(a) $\geq$ 100

```{r}
pnorm(100 - 100, mean = 0, sd = sqrt(364/4), lower.tail = FALSE)
```

(b) $\geq$ 110

```{r}
pnorm(110 - 100, mean = 0, sd = sqrt(364/4), lower.tail = FALSE)
```

(c) $\geq$ 120

```{r}
pnorm(120 - 100, mean = 0, sd = sqrt(364/4), lower.tail = FALSE)
```


### Calculate the expected value and variance of the binomial distribution using the moment generating function.

Moment generating function for a random variable X: $M(t)=[(1-p) +pe^t]^n$

First Derivative $M(t)$: $M'(t) = n(pe^t)[(1-p)+pe^t]^{n-1}$

for t = 0: $M'(0) = n [ 1 - p + p ]^{n-1} p(1) = np$

Expected value: $E(X) = \mu = M'(0) = np$

Variance: $M''(t) = n[1 - p + pe^t]^{n-1}(pe^t)+ (pe^t) n(n-1)[1 - p + pe^t]^{n-2}(pe^t)$

for t = 0: $M''(0) = np + n^2p^2 -np^2$

Variance: $M''(0) = n(n-1)p^2 + np$ OR $\sigma^2 = np(1-p)$


### Calculate the expected value and variance of the exponential distribution using the moment generating function.

Moment generating function for exponential distribution: $f(x)= \lambda e^{−\lambda x}$

$$g(t)=∫_{ 0 }^{ \infty }e^{ tx }λe^{ -λe }dx$$

$$g(t)=\frac { λe^{ (t-λ)x } }{ t-λ } |_{ 0 }^{ ∞ }$$

$$g(t)=\frac { λ }{ λ-t }$$

$$g(t)=\frac { λ }{ (λ-t)^{ 2 } }$$

$$g(t)=\frac { λ }{ λ^{ 2 } } =\frac { 1 }{ λ }$$

$$g''(t)=\frac { 2λ }{ (λ-t)^{ 3 } }$$

$$g''(0)=\frac { 2λ }{ λ^{ 3 } } =\frac { 2 }{ λ^{ 2 } }$$

Expected Value:

$$\mu =g'(0)=λ^{ -1 }$$

Variance:

$$σ^{ 2 }=g''(0)-g'(0)^{ 2 }=\frac { 2 }{ λ^{ 2 } } =\frac { 1 }{ λ^{ 2 } } =λ^{ -2 }$$

