---
title: "LAW OF LARGE NUMBERS - Assignment 8"
author: "Monu Chacko"
date: "3/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 11. A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?

$$E[X_i] = \frac{1}{\lambda_i} = 1000$$

#### Expected lifetime of a bulb is 1000 hours.

$$\lambda_i = \frac{1}{1000}$$

$X_i$ is exponential so

$$min\{X_1,X_2,...,X_{100}\} \sim exponential(\sum\limits_{i=1}^{100} \lambda_i)$$

$$E[min X_i] = \frac{1}{\frac{1}{10}} = 10$$


#### Expected time for the ﬁrst of these bulbs to burn out: 10 hours




### 14. Assume that X1 and X2 are independent random variables, each having an exponential density with parameter $\lambda$. Show that $Z = X_1 − X_2$ has density

$$f_Z(z)=(1/2)\lambda e^{−\lambda|z|}$$


#### Probability Density Function

$$f(x_1) = \lambda e^{-\lambda x_1}$$

$$f(x_2) = \lambda e^{-\lambda x_2}$$


#### Joint density

$$\lambda^2 e^{-\lambda(x_1 + x_2)}$$

Substitute to get joint density $Z$ and $X_2$ 
$$\lambda^2 e^{-\lambda(z + 2x_2)}$$

When z is positive:

$$\int_{0}^{\infty} \lambda^2 e^{-\lambda(z + 2x_2)} dx = \frac{\lambda}{2} e^{-\lambda z}$$

When z is negative:

$$\int_{-z}^{\infty} \lambda^2 e^{-\lambda(z + 2x_2)} dx = \frac{\lambda}{2} e^{\lambda z}$$


$$f_Z(z) = \frac{1}{2} \lambda e^{\lambda|z|}$$

### 1. Let X be a continuous random variable with mean $\mu = 10$ and variance  $\sigma^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.


```{r}
chebyshev <- function(e, var){
    pX = var / e^2
    return(pX)
}
```

(a) $P(|X−10| \geq 2)$

```{r}
var <- 100/3
e <- 2
p <- chebyshev(e, var)
p
```

(b) $P(|X−10| \geq 5)$

```{r}
e <- 5
p <- chebyshev(e, var)
p
```

(c) $P(|X−10| \geq 9)$

```{r}
e <- 9
p <- chebyshev(e, var)
p
```

(e) $P(|X − 10| \geq 20)$

```{r}
e <- 20
p <- chebyshev(e, var)
p
```

