---
title: "FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS"
author: "Monu Chacko"
date: "2/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(matrixcalc)
library(pracma)
```

## ASSIGNMENT 3

1.Problem set 1

(1)  What is the rank of the matrix A?

$$
A=
\left[
  \begin{array}{cccc}
  1 & 2 & 3 & 4 \\
  -1 & 0 & 1 & 3 \\
  0 & 1 & -2 & 1 \\
  5 & 4 & -2 & -3
  \end{array}
\right]
$$

```{r}
A <- matrix(c(1, 2, 3, 4,-1, 0, 1, 3, 0, 1, -2, 1, 5, 4, -2, -3), nrow=4, byrow = TRUE)
A

det(A)

# Since the determinant is non-zero the rank is:
Rank(A)
```

(2)  Given an mxn matrix where m > n, what can be the maximum rank?  The minimum rank, assuming that the matrix is non-zero?

Maximum the rank would be the number of rows or number of columns (if you transpose). Rank is the total number of pivot points. The minimum is 1 since it is non-zero.


(3)  What is the rank of matrix B?

$$
B=
\left[
  \begin{array}{cccc}
  1 & 2 & 1 \\
  3 & 6 & 3 \\
  2 & 4 & 2
  \end{array}
\right]
$$


```{r}
B <- matrix(c(1,2,1,3,6,3,2,4,2), nrow=3, byrow=TRUE)
B
```


If we reduce this to Row Echelon form, we can see the pivot of 1. So the rank is 1. Row 2 and 3 are multiples of Row 1. So it makes sense.

```{r}
rref(B)
matrix.rank(B)
```

2.Problem set 2


Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work.  You'll need to write out the characteristic polynomial and show your solution.

$$
A=
\left[
  \begin{array}{cccc}
  1 & 2 & 3 \\
  0 & 4 & 5 \\
  0 & 0 & 6
  \end{array}
\right]
$$

This is a nxn (square) matrix and is also an upper triangular matrix in which all entries below the main diagonal are zero. 

The eigenvalues are values on the diagonal i.e 1,4,6.  

$$\lambda_1=1$$
$$\lambda_2=4$$
$$\lambda_3=6$$


```{r}
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow=3, byrow=TRUE)
A
eigen(A)$values
```

Characteristic Polynomial is Det(A-位I)=0 or Det(位I-A)=0 where I is the identity matrix  

```{r}
# Test for Characteristic Polynomial
det(A - (rref(A) * diag(x=1, nrow=3))) == 0

```

$$f_A(\lambda) = (1-\lambda)(4-\lambda)(6-\lambda) = 24-34\lambda+11\lambda^2-\lambda^3$$

If $\lambda=1$, then $A - 1I_3$ is row-reduced to

$$
\begin{bmatrix}
 0 &1 &0\\
 0 &0 &1\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
 $$

Eigenspace would be:

$$
E_{\lambda=1}=
 span
 \Bigg(
 \begin{bmatrix}
 1\\
 0\\
 0
 \end{bmatrix}
  \Bigg)
$$


If 位=4

$$
\begin{bmatrix}
 1 &-\frac{2}{3} &0\\
 0 &0 &1\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$ 


Eigenspace would be:  

$$
E_{\lambda=4}=
 span
 \Bigg(
 \begin{bmatrix}
 1\\
 1.5\\
 0
 \end{bmatrix}
 \Bigg)
$$


 if 位=6
 
 $$
 \begin{bmatrix}
 1 &0 &-1.6\\
 0 &1 &-2.5\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$

Eigenspace would be:  

$$
E_{\lambda=6}=
 span
 \Bigg(
 \begin{bmatrix}
 1.6\\
 2.5\\
 1
 \end{bmatrix}
 \Bigg)
$$