---
title: "CleanUp"
author: "Monu Chacko, Abdelmalek Hajjam, Md Forhad Akbar, Shovan Biswas"
date: "10/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE, message = FALSE}

library(caret)
library(skimr)
library(data.table)
library(mltools)
library(tidyverse)
library(corrplot)
library(ROCR)
library(pROC)
library(DMwR)
library(Rtsne)
library(glmnet)
library(doParallel)
library(gridExtra)
theme_set(theme_bw())
```
## Reading Original mydata 

```{r}
# Reading in data
mydata <- read.csv("diabetic_data.csv")
# remove duplicates
mydata <- mydata[!duplicated(mydata$patient_nbr), ]
```

## Preparing and processing
```{r}
mydata[mydata == "?"] <- NA
```

Dropping unnecessary/unvaluable variables or variables with too many missing values    
```{r}
drops <- c("encounter_id", "patient_nbr", "weight", "payer_code", 
           "medical_specialty", "diabetesMed", "diag_2", "diag_3")
mydata <- mydata[ , !(names(mydata) %in% drops)]
```

any patient who's discharge status is "expired" will be dropped.
```{r}
mydata <- filter(mydata, !(discharge_disposition_id %in% c(11, 19, 20, 21)))
```

Remove unknown gender
```{r}
mydata <- mydata %>% filter(gender != "Unknown/Invalid")
mydata$gender <- droplevels(mydata$gender)
```

## Let's Encode
```{r}
#encoding gender Female=0, Male=1
mydata$gender <- as.integer(mydata$gender) - 1
```

```{r}
# Encoding ages - age will be the median of the 10 years age interval.
mydata$age <- ifelse(mydata$age == "[0-10)", 5, 
                  ifelse(mydata$age == "[10-20)", 15, 
                         ifelse(mydata$age == "[20-30)", 25, 
                                ifelse(mydata$age == "[30-40)", 35, 
                                       ifelse(mydata$age == "[40-50)", 45, 
                                              ifelse(mydata$age == "[50-60)", 55, 
                                                     ifelse(mydata$age == "[60-70)", 65, 
                                                            ifelse(mydata$age == "[70-80)", 75, 
                                                                   ifelse(mydata$age == "[80-90)", 85, 95)))))))))
                                                                    
```

```{r}
#encoding Glucose - measurment = 0; Normal reading = 1; >200 & >300 (abnormal readings) = 2
mydata$max_glu_serum <- ifelse(mydata$max_glu_serum == "None", 0, 
                          ifelse(mydata$max_glu_serum == "Norm", 1, 2))
```

```{r}
#encoding A1C test results - measurment = 0; Normal reading = 1; abnormal reading (>7 or >8) = 2
mydata$A1Cresult <- ifelse(mydata$A1Cresult == "None", 0, 
                         ifelse(mydata$A1Cresult == "Norm", 1, 
                                ifelse(mydata$A1Cresult %in% c(">7", ">8"), 2, NA)))
```

```{r}
#encoding insulin dosage - no insulin = 0; decrease in insulin = 1; steady insulin = 2; increase in insulin = 3
mydata$insulin <- ifelse(mydata$insulin == "No", 0, ifelse(mydata$insulin == "Down", 1, ifelse(mydata$insulin == "Steady", 2, 3)))
```

```{r}
#encoding Medication Change - no change in medication = 0; change in medication = 1
mydata$change <- ifelse(mydata$change == "Ch", 1, 0)
```

```{r}
#encoding race - Caucasian = 0; African American = 1; Other = 2
mydata$race <- ifelse(mydata$race == "Caucasian", 0, ifelse(mydata$race == "AfricanAmerican", 1, 2))
mydata$race[is.na(mydata$race)] <- 2
```

```{r}
#encoding target variable "readmitted" - No Readmission within 30 days = 0; 1 for a Readmission in <30 days = 1
mydata$readmitted <- ifelse((mydata$readmitted == "NO" | mydata$readmitted == ">30"), 0, 1)
mydata %>% group_by(readmitted) %>% summarize(count=n())
```

## combining some features
```{r}
# Sum of number of outpatient, emergency and inpatient encounters 
mydata$healthcare_utilization <- (mydata$number_outpatient + mydata$number_emergency + mydata$number_inpatient)
```

```{r}
#sum of other non insulin medications
mydata$metformin <- ifelse(mydata$metformin %in% c("Up", "Steady", "Down"),1,0)
mydata$repaglinide <- ifelse(mydata$repaglinide %in% c("Up", "Steady", "Down"),1,0)
mydata$nateglinide <- ifelse(mydata$nateglinide %in% c("Up", "Steady", "Down"),1,0)
mydata$chlorpropamide <- ifelse(mydata$chlorpropamide %in% c("Up", "Steady", "Down"),1,0)
mydata$glimepiride <- ifelse(mydata$glimepiride %in% c("Up", "Steady", "Down"),1,0)
mydata$acetohexamide <- ifelse(mydata$acetohexamide %in% c("Up", "Steady", "Down"),1,0)
mydata$glipizide <- ifelse(mydata$glipizide %in% c("Up", "Steady", "Down"),1,0)
mydata$glyburide <- ifelse(mydata$glyburide %in% c("Up", "Steady", "Down"),1,0)
mydata$tolbutamide <- ifelse(mydata$tolbutamide %in% c("Up", "Steady", "Down"),1,0)
mydata$pioglitazone <- ifelse(mydata$pioglitazone %in% c("Up", "Steady", "Down"),1,0)
mydata$rosiglitazone <- ifelse(mydata$rosiglitazone %in% c("Up", "Steady", "Down"),1,0)
mydata$acarbose <- ifelse(mydata$acarbose %in% c("Up", "Steady", "Down"),1,0)
mydata$miglitol <- ifelse(mydata$miglitol %in% c("Up", "Steady", "Down"),1,0)
mydata$troglitazone <- ifelse(mydata$troglitazone %in% c("Up", "Steady", "Down"),1,0)
mydata$tolazamide <- ifelse(mydata$tolazamide %in% c("Up", "Steady", "Down"),1,0)
mydata$examide <- ifelse(mydata$examide %in% c("Up", "Steady", "Down"),1,0)
mydata$citoglipton <- ifelse(mydata$citoglipton %in% c("Up", "Steady", "Down"),1,0)
mydata$glyburide.metformin <- ifelse(mydata$glyburide.metformin %in% c("Up", "Steady", "Down"),1,0)
mydata$glipizide.metformin <- ifelse(mydata$glipizide.metformin %in% c("Up", "Steady", "Down"),1,0)
mydata$glimepiride.pioglitazone <- ifelse(mydata$glimepiride.pioglitazone %in% c("Up", "Steady", "Down"),1,0)
mydata$metformin.rosiglitazone <- ifelse(mydata$metformin.rosiglitazone %in% c("Up", "Steady", "Down"),1,0)
mydata$metformin.pioglitazone <- ifelse(mydata$metformin.pioglitazone %in% c("Up", "Steady", "Down"),1,0)
mydata$num_other_diabetes_meds_up_stdy_dwn <- (mydata$metformin + mydata$repaglinide + mydata$nateglinide + mydata$chlorpropamide + mydata$glimepiride + mydata$acetohexamide + mydata$glipizide + mydata$glyburide + mydata$tolbutamide + mydata$pioglitazone + mydata$rosiglitazone + mydata$acarbose + mydata$miglitol + mydata$troglitazone + mydata$tolazamide + mydata$examide + mydata$citoglipton + mydata$glyburide.metformin + mydata$glipizide.metformin + mydata$glimepiride.pioglitazone + mydata$metformin.rosiglitazone + mydata$metformin.pioglitazone)
drops2 <- c("metformin", "repaglinide", "nateglinide", "chlorpropamide", "glimepiride", "acetohexamide", "glipizide", 
            "glyburide", "tolbutamide", "pioglitazone", "rosiglitazone", "acarbose", "miglitol", "troglitazone", 
            "tolazamide", "examide", "citoglipton", "glyburide.metformin", "glipizide.metformin", "glimepiride.pioglitazone",
            "metformin.rosiglitazone", "metformin.pioglitazone")
mydata <- mydata[ , !(names(mydata) %in% drops2)]
```

## One Hot encoding
```{r}
#dummy variables for admission types
mydata <- mutate(mydata, at_emergent = as.numeric(admission_type_id %in% c(1, 2, 7)), 
              at_elective = as.numeric(admission_type_id == 3), 
              at_other = as.numeric(admission_type_id %in% c(4, 5, 6, 8)))

#dummy variables for discharge dispositions
mydata <- mutate(mydata, dd_home = as.numeric(discharge_disposition_id %in% c(1, 6, 8)), 
              dd_facility_transfer = as.numeric(discharge_disposition_id %in% c(2, 3, 4, 5, 10, 16, 17, 22, 23, 24, 30, 27, 28, 29, 13, 14)), 
              dd_other = as.numeric(discharge_disposition_id %in% c(7, 18, 25, 26, 9, 12, 15))) 
             
#dummy variables for admission source
mydata <- mutate(mydata, as_outpatient = as.numeric(admission_source_id %in% c(1, 2, 3)),
              as_facility_transfer = as.numeric(admission_source_id %in% c(4, 5, 6, 10, 18, 19, 22, 25, 26)),
              as_ed = as.numeric(admission_source_id %in% c(7)),
              as_other = as.numeric(admission_source_id %in% c(8, 9, 15, 17, 20, 21, 11, 12, 13, 14, 23, 24)))
              
mydata <- within(mydata, rm(admission_type_id))
mydata <- within(mydata, rm(discharge_disposition_id))
mydata <- within(mydata, rm(admission_source_id))
```

```{r}
#grouping the ICD codes into 20 different categories based on the Strack et al. 2014 research paper
mydata$diag_1 <- as.character(mydata$diag_1)
mydata$diag_1grp <- ifelse (mydata$diag_1 == "?", "Unknown", 
          ifelse(grepl(mydata$diag_1, pattern = "[EV]") == T, "External", 
            ifelse(floor(as.numeric(mydata$diag_1)) == 250, "Circulatory", 
              ifelse(mydata$diag_1 %in% c(390:459, 785), "Diabetes", 
                ifelse(mydata$diag_1 %in% c(460:519, 786), "Respiratory", 
                  ifelse(mydata$diag_1 %in% c(520:579, 787), "Digestive", 
                    ifelse(mydata$diag_1 %in% c(800:999), "Injury", 
                      ifelse(mydata$diag_1 %in% c(710:739), "Musculoskeletal",
                        ifelse(mydata$diag_1 %in% c(580:629, 788), "Genitourinary", 
                         ifelse(mydata$diag_1 %in% c(140:239), "Neoplasm", 
                           ifelse(mydata$diag_1 %in% c(780, 781, 784, 790:799, 740:759), "Other", # added congenital to other
                             ifelse(mydata$diag_1 %in% c(240:249, 251:279), "Endocrine_Nutrition_Metabolic",  
                               ifelse(mydata$diag_1 %in% c(680:709, 782), "Skin", 
                                 ifelse(mydata$diag_1 %in% c(001:139), "Infectious",
                                   ifelse(mydata$diag_1 %in% c(290:319), "Mental", 
                                     ifelse(mydata$diag_1 %in% c(280:289), "Blood",
                                       ifelse(mydata$diag_1 %in% c(320:359), "Nervous",
                                         ifelse(mydata$diag_1 %in% c(630:679), "Pregnancy", 
                                          ifelse(mydata$diag_1 %in% c(360:389), "Sense", "Unknown") 
                                ))))))))))))))))))
mydata$diag_1grp <- as.factor(mydata$diag_1grp)
mydata <- one_hot(data.table(mydata))
mydata <- as.data.frame(mydata) # convert back to data.frame format
mydata <- within(mydata, rm(diag_1))
mydata <- na.omit(mydata)
```

## Removing Highly Correlated Variables
We will remove any highly correlated variables by investigating the correlation matrix
```{r, fig.height=15, eval=F, echo=F}
# moving readmitted to the last column value
mydata <- mydata %>% select(-readmitted, readmitted) 
cor_mat <- cor(mydata[, 1:42])
plot.new()
corrplot(cor_mat, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45,mar=c(1,1,1,1))
```

```{r correlation matrix, fig.width=8, fig.height=10}
mydata <- mydata %>% select(-readmitted, readmitted) # move readmitted to the last column value
cor_mat <- cor(mydata[, 1:42])
cor_mat[!lower.tri(cor_mat)] <- 0
mydata <- mydata[,!apply(cor_mat,2,function(x) any(abs(x) > 0.6))]
cor_mat2 <- cor(mydata[, 1:42])
get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
}
cor_mat_melted <- melt(get_lower_tri(cor_mat2), na.rm = T)
ggplot(data = cor_mat_melted, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = -45, vjust = 1, hjust = 0),
        legend.position = c(0,1), legend.justification = c(0,1),
        axis.title = element_blank(), plot.title = element_text(hjust = 0.5)) +
  labs(title = "Correlation matrix") +
  scale_y_discrete(position = "right")
```

## Obtain a cleaned dataset we can work with
```{r}
# write the cleaned mydata to a new file
write.csv(mydata, file = "diabetic_cleanedANDprocessed.csv", row.names = F)
```


