---
title: "DATA 624 Homework 10 - Market Basket Analysis"
author: "Monu Chacko"
date: "5/9/2021"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Imagine 10000 receipts sitting on your table. Each receipt represents a transaction with items that were purchased. The receipt is a representation of stuff that went into a customer’s basket - and therefore ‘Market Basket Analysis’.

That is exactly what the Groceries Data Set contains: a collection of receipts with each line representing 1 receipt and the items purchased. Each line is called a transaction and each column in a row represents an item.  The data set is attached.

Your assignment is to use R to mine the data for association rules. You should report support, confidence and lift and your top 10 rules by lift. 

Do a simple cluster analysis on the data as well.  Use whichever packages you like.


```{r warning=FALSE, message=FALSE}
library(arules)
library(RColorBrewer)
library(kableExtra)
```


## Load Data

***Let's load the data and examine it***

```{r warning=FALSE, message=FALSE}
data <- read.csv("https://raw.githubusercontent.com/monuchacko/cuny_msds/master/data_624/data/GroceryDataSet.csv", header=FALSE)

# View Summary
summary(data) %>%
  kable() %>%
    kable_styling()
```

```{r warning=FALSE, message=FALSE}
groceryDataset = read.transactions("https://raw.githubusercontent.com/monuchacko/cuny_msds/master/data_624/data/GroceryDataSet.csv", sep = ',', rm.duplicates = TRUE)
```

## Analyse data

```{r warning=FALSE, message=FALSE}
# View summary
summary(groceryDataset)
```

***As we see from the summary, 'whole milk' is the most frequent item with 2513 and then followed by 'other vegetables' with 1903. Lets see a visual using the item frequency plot.*** 


```{r warning=FALSE, message=FALSE}
itemFrequencyPlot(groceryDataset,topN=10, type="absolute", col=brewer.pal(8,'Pastel2'), main="Absolute Item Frequency Plot")
```

***We see a better visual using the frequency plot above. ItemFrequencyPlot was used to create an item Frequency Bar Plot to view the distribution of objects based on itemMatrix.***

## Train/ Extract

***Lets train the apriori to extract rules by defining the minimim support and confidence value. This is basically the likelihood of the product to be purchased.***

```{r warning=FALSE, message=FALSE}
min_suport <- 6 * 7/ nrow(groceryDataset)
min_suport
```

***We find the confidence which is the likelihood of a product being purchased given another product is purchased.***

***Confidence(p1 -> p2) = # of observation where p1 and p2 purchased/ # of observation where p1 purchased***

```{r warning=FALSE, message=FALSE}
# Training Apriori on the grocery dataset
rules = apriori(data = groceryDataset, parameter = list(support = 0.004, confidence = 0.6))
```

***Inspect top 10 rules by lift. Lift indicates the significance of the rule.***

```{r warning=FALSE, message=FALSE}
inspect(sort(rules, by = 'lift')[1:10])
```

***We started with 0.6 confidence. Lets reduce the confidence to 0.4 and see if it is better.***

```{r warning=FALSE, message=FALSE}
# Training Apriori on the dataset
rules = apriori(data = groceryDataset, parameter = list(support = 0.004, confidence = 0.4))
```

***Let see the performance after we changed the confidence to 0.4***

```{r warning=FALSE, message=FALSE}
# Visualising the results
inspect(sort(rules, by = 'lift')[1:10])
```

***Although it looks better, there is more room for improvement. For example we see 'citrus fruit' in multiple rules. We can further change the minimum value to 0.2 and evaluate its performance.***

```{r warning=FALSE, message=FALSE}
# Training Apriori on the dataset
rules = apriori(data = groceryDataset, parameter = list(support = 0.004, confidence = 0.2))
```


```{r warning=FALSE, message=FALSE}
# Visualising the results
inspect(sort(rules, by = 'lift')[1:10])
```

***As we see we have better results and the association rules looks better.***
