---
title: "HW10"
author: "Monu Chacko"
date: "5/9/2021"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Imagine 10000 receipts sitting on your table. Each receipt represents a transaction with items that were purchased. The receipt is a representation of stuff that went into a customer’s basket - and therefore ‘Market Basket Analysis’.

That is exactly what the Groceries Data Set contains: a collection of receipts with each line representing 1 receipt and the items purchased. Each line is called a transaction and each column in a row represents an item.  The data set is attached.

Your assignment is to use R to mine the data for association rules. You should report support, confidence and lift and your top 10 rules by lift. 

Extra credit: do a simple cluster analysis on the data as well.  Use whichever packages you like.  Due May 3 before midnight.

```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```

