---
title: "DATA 624 Homework 3 - Time series decomposition"
author: "Monu Chacko"
date: "2/28/2021"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r warning=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
library(ggplot2)
library(fpp2)
library(seasonal)
library(readxl)
```

## Question 6.2

The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.

a. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?

```{r}
summary(plastics)
```

```{r}
plastics
```

```{r}
autoplot(plastics) + xlab("Month") + ylab("Sales (in thousands)") +
  ggtitle("Monthly sales (in thousands) of product A for a plastics manufacturer for five years")
```


```{r}
ggseasonplot(plastics)
```


```{r}
ggsubseriesplot(plastics)
```


```{r}
gglagplot(plastics)
```


```{r}
Acf(plastics)
```


The trends show that there is seasonality in the sales for plastic A. Looks like the sales increases from spring through summer. During fall and winter there is a decline in sales. Overall, there is a general increased positive trend year over year.

```{r}
plastics %>%
  stl(t.window=12, s.window="periodic", robust=TRUE) %>%
  autoplot()
```

b.Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

```{r}
multiplicative_decomp <- decompose(plastics, type = "multiplicative")
multiplicative_decomp
```

```{r}
autoplot(multiplicative_decomp) + xlab("Month") +
  ggtitle("Monthly sales (in thousands) of product A for a plastics manufacturer for five years")
```

c.Do the results support the graphical interpretation from part A

Yes, the graphical interpretation supports the findings in part A

d.Compute and plot the seasonally adjusted data.

```{r}
autoplot(plastics, series="Data") +
  autolayer(trendcycle(multiplicative_decomp), series="Trend") +
  autolayer(seasadj(multiplicative_decomp), series="Seasonally Adjusted") +
  xlab("Year") + ylab("Monthly Sales") +
  ggtitle("Plastic Sales") +
  scale_colour_manual(values=c("green","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))
```

The trend line is upward trending

e.Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

```{r}
change.plastics.2 <- plastics
change.plastics.2[2] <- plastics[2]+500
change.plastics.2
```


```{r}
change_multiplicative_decomp <- decompose(change.plastics.2, type = "multiplicative")
autoplot(change.plastics.2, series="Data") +
  autolayer(trendcycle(change_multiplicative_decomp), series="Trend") +
  autolayer(seasadj(change_multiplicative_decomp), series="Seasonally Adjusted") +
  xlab("Year") + ylab("Monthly Sales") +
  ggtitle("Plastic Sales") +
  scale_colour_manual(values=c("green","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))
```

This changes the Seasonally Adjusted.

f.Does it make any difference if the outlier is near the end rather than in the middle of the time series?

```{r}
change.plastics.middle <- plastics
change.plastics.middle[25] <- plastics[25]+500

change.plastics.end <- plastics
change.plastics.end[59] <- plastics[59]+500

change_multiplicative_decomp_middle <- decompose(change.plastics.middle, type = "multiplicative")
autoplot(change.plastics.middle, series="Data") +
  autolayer(trendcycle(change_multiplicative_decomp_middle), series="Trend") +
  autolayer(seasadj(change_multiplicative_decomp_middle), series="Seasonally Adjusted") +
  xlab("Year") + ylab("Monthly Sales") +
  ggtitle("Plastic Sales - Outlier in the Middle") +
  scale_colour_manual(values=c("green","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))
```


```{r}
change_multiplicative_decomp_end <- decompose(change.plastics.end, type = "multiplicative")
autoplot(change.plastics.end, series="Data") +
  autolayer(trendcycle(change_multiplicative_decomp_end), series="Trend") +
  autolayer(seasadj(change_multiplicative_decomp_end), series="Seasonally Adjusted") +
  xlab("Year") + ylab("Monthly Sales") +
  ggtitle("Plastic Sales - outlier at the end") +
  scale_colour_manual(values=c("green","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))
```

The seasonality is  adjusted and the trend had a effect on the outliers at the middle and end.

## Question 6.3

Recall your retail time series data (from Exercise 3 in Section 2.10). Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?

```{r}
retail_data <- read_excel("data/retail.xlsx", skip = 1)
retail <- ts(retail_data[, "A3349397X"], frequency = 12, start = c(1982, 4))
x11_retail <- seas(retail, x11 = "")
autoplot(x11_retail) + ggtitle("X11 Decomposition of Retail Sales")
```

The remainder shows outliers around 2000-2001 which may not have been previously noticed.

