---
title: "Project 2"
author: "Abdelmalek, Monu Chacko, Paul Perez"
date: "5/22/2021"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Problem Statement

***TODO:***

This is role playing. I am your new boss. I am in charge of production at ABC Beverage and you are a team of data scientists reporting to me. My leadership has told me that new regulations are requiring us to understand our manufacturing process, the predictive factors and be able to report to them our predictive model of PH. Please use the historical data set I am providing. Build and report the factors in BOTH a technical and non-technical report. I like to use Word and Excel. Please provide your non-technical report in a business friendly readable document and your predictions in an Excel readable format. The technical report should show clearly the models you tested and how you selected your final approach. Please submit both Rpubs links and .rmd files or other readable formats for technical and non-technical reports. Also submit the excel file showing the prediction of your models for pH.

***TODO:***

# Executive Summary

***TODO:***
New Regulations by ABC beverage company leadership requires the company’s production unit to better understand the manufacturing process, the predictive factors and their relationship to the PH of the beverages.
***TODO:***

# Research Statement

***TODO:***
The research is an effort to find the predictive variables related to the ph of the beverages and build the predictive model for ph of beverages
***TODO:***

# Data Collection

***TODO:***
The data set is a historic data containing predictors associated to the PH and is provided in an excel file. We will utilize this historic dataset to analyze and predict the PH of beverages. Two excel files are provided:

- The training data (StudentData.xlsx)
- The test data (StudentEvaluation.xlsx).
***TODO:***

```{r warning=FALSE, message=FALSE}
library("readxl")
library(httr)
library(caret)
library(tidyverse)
library(kableExtra)
library(caretEnsemble)
library(mice)
library(kableExtra)
library(xgboost)
library(dplyr)
library(parallel)
library(doParallel)
library(ggplot2)
library(Hmisc)
library(psych)
library(reshape2)
library(gridExtra)
library(rpart.plot)
library(DT)
```


```{r warning=FALSE, message=FALSE}
se_data <- read.csv("https://raw.githubusercontent.com/monuchacko/cuny_msds/master/data_624/data/StudEval.csv")
sd_data <- read.csv("https://raw.githubusercontent.com/monuchacko/cuny_msds/master/data_624/data/StudData.csv")

```

```{r warning=FALSE, message=FALSE}
dim(sd_data)
```

```{r warning=FALSE, message=FALSE}
colnames(sd_data)
```

```{r warning=FALSE, message=FALSE}
str(sd_data)
```

```{r warning=FALSE, message=FALSE}
dim(se_data)
```

```{r warning=FALSE, message=FALSE}
colnames(se_data)
```

```{r warning=FALSE, message=FALSE}
str(se_data)
```




```{r warning=FALSE, message=FALSE}

#***TODO:***

#***TODO:***

bev_raw <- read_csv('https://raw.githubusercontent.com/hovig/Team5-Data624-Project2/master/StudentData.csv') 
head(bev_raw)
```

# Data Preparation and EDA (Exploratory Data Analysis)

***TODO:***

***TODO:***


```{r warning=FALSE, message=FALSE}
bev_raw %>%
  ggplot(aes(PH, fill=PH > 9)) + 
  geom_histogram(bins=30) +
  theme_bw() +
  theme(legend.position='center') +
  labs(y='Count', title='PH Levels in Dataset')
```



```{r warning=FALSE, message=FALSE}
bev_raw <- bev_raw %>% 
  filter(!is.na(bev_raw$PH), bev_raw$PH < 9) 
```

```{r}
dim(bev_raw)
str(bev_raw)
hist.data.frame(bev_raw)
table(bev_raw$`Brand Code`)
summary(bev_raw)
describe(bev_raw %>% select(-`Brand Code`))
```

## Zero variance

***TODO:***
To filter for near-zero variance predictors, the caret package function nearZeroVar will return the column numbers of any predictors that fulfill the conditions outlined. A zero variance predictor will never be chosen for a split since it offers no possible predictive information.
***TODO:***

```{r warning=FALSE, message=FALSE}
df1 <- bev_raw %>% select(-`Brand Code`) %>% mutate_each(funs(as.numeric(.)))%>%complete.cases()%>%
       as.data.frame()


names(df1)[nearZeroVar(df1)]
```


```{r warning=FALSE, message=FALSE}
nzv <- nearZeroVar(bev_raw,saveMetrics= TRUE)
nzv[nzv$nzv,]
```

## Box plot

***TODO:***
Box plots for the variables reveal, that besides having the outliers in the variables, most of the variables are skewed. For example: Variables density, carb flow, filler speed and oxygen filler are skewed providing us an opportunity to further check their distribution.
***TODO:***



```{r warning=FALSE, message=FALSE}
df.m <- melt(bev_raw %>% select(-MFR, -`Filler Speed`, -`Carb Flow`,-`Bowl Setpoint`,`Carb Pressure1`,
            `Hyd Pressure4`, `Air Pressurer`, `Carb Temp`, `Filler Level`, `Mnf Flow`), id.var = "Brand Code")
p <-ggplot(data = df.m, aes(x=variable, y=value)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4,aes(fill=variable)) +     
  scale_y_continuous(name = "Predictors for PH", breaks = seq(0, 2, 0.5))  + 
  coord_flip()
p
```


```{r warning=FALSE, message=FALSE}

```




## Normality

***TODO:***
Normality is one of the most widely used technique to understand the continuous predictors. In the below plot we can see normal distribution behavior of the given dataset for different features.
***TODO:***

```{r warning=FALSE, message=FALSE}
bev_raw%>%
 select(-`Brand Code`) %>%   
  select(2:20) %>%         
  gather() %>%                            
  ggplot(aes(value)) +                     
    facet_wrap(~ key, scales = "free") +  
    geom_density() 
```



```{r warning=FALSE, message=FALSE}
bev_raw%>%
 select(-`Brand Code`) %>%   
  select(`Carb Flow`, Density, MFR, Balling, `Pressure Vacuum`, PH, `Oxygen Filler`, `Bowl Setpoint`, `Pressure Setpoint`, `Air Pressurer`, `Alch Rel`,`Carb Rel`,`Balling Lvl`)   %>%               
  gather() %>%                             
  ggplot(aes(value)) +                     
    facet_wrap(~ key, scales = "free") +  
    geom_density() 
```



# Data Preprocessing

***TODO:***
We perform 3 data preparation steps - Remove the near-zero variance variables we previously mentioned - Impute missing values with a Random Forest Regression with the MICE package - Create dummy variables for the categorical variable, brand

Random Forest was chosen as the regression method in imputation because it requires very little tuning, allowing and there is no interface to tune the imputation model in MICE.
***TODO:***


```{r warning=FALSE, message=FALSE}
cl <- makeCluster(detectCores())
registerDoParallel(cl)
set.seed(42)

bev <- select(bev_raw, -PH) %>%
  union_all(se_data) #since we're not using the labels (y values), combining the sets for data clearning is fine
cols <- str_replace(colnames(bev), '\\s', '')
colnames(bev) <- cols
bev <- mutate(bev, BrandCode = ifelse(is.na(BrandCode), 'Unknown', BrandCode))
y <- bev_raw$PH

zero_vars <- nearZeroVar(bev[1:2566, ])
bev_new <- bev[, -zero_vars]

pred <- mice(data = bev_new, m = 3, method = 'rf', maxit = 3)
bev_imputed <- complete(pred)

form <- dummyVars(~ ., data = bev_imputed)
bev_imputed <- predict(form, bev_imputed) %>% data.frame() %>% as.matrix()
```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


```{r warning=FALSE, message=FALSE}

```


# Model Building

***TODO:***

***TODO:***

## Test Train Split

***TODO:***

***TODO:***


## Modeling Technique and Approach

***TODO:***

***TODO:***


## MARS

***TODO:***

***TODO:***


## RandomForest

***TODO:***

***TODO:***


## Cubist

***TODO:***

***TODO:***


## XGB Trees

***TODO:***

***TODO:***


## XGB Dart

***TODO:***

***TODO:***


# Model Evaluation And Model Selection

***TODO:***

***TODO:***


## Variable Importances

***TODO:***

***TODO:***


## Test Model Stack

***TODO:***

***TODO:***


## Model Selection

***TODO:***

***TODO:***


## Predict on Eval Set

***TODO:***

***TODO:***


# Conclusion

***TODO:***

***TODO:***


# References

***TODO:***

***TODO:***





